RIScatter: Unifying Backscatter Communication and Reconfigurable Intelligent Surface

### 2023-07-31 review/response.tex

	\point The link between eq. (9) and eq. (25) lacks clarity and misses intermediate steps. Also, the expression of the function $Q$ provided in eq. (25) appears to be independent of the active beamforming $\boldsymbol{w}$.  It is unclear how the derivation of $Q$ with respect to $\boldsymbol{w}$ is performed in this case. Please clarify the relationship between $Q$ and $\boldsymbol{w}$ and provide a clear derivation for better understanding. The same remark about clarity applies to eq. (28) and eq. (33).
	\reply {$Q\Bigl(N,\frac{t_{l-1}}{\sigma_{m_{\mathcal{K}}}^2},\frac{t_l}{\sigma_{m_{\mathcal{K}}}^2}\Bigr)$ is the regularized incomplete Gamma function that depends on the variance of accumulated receive energy $\sigma_{m_{\mathcal{K}}}^2$, which is a function of $\boldsymbol{w}$. We have added a reminder on this, and have added the intermediate steps from (9) to (25) with proper details. The corresponding part has been completely rewritten to improve logic and clarity.}

	\point Is the BCD algorithm convergent for the optimization problem (20)?
	\reply {The BCD algorithm is convergent for the optimization problem (20), since the input distribution and active beamforming subproblems converge and the thresholding subproblem attains global optimality. This argument has been added prior to the convergence results in Fig. 7.}

	\point Please write the mathematical expression of eq. (40) properly. Moreover, there appears to be a mistake in the sign of the Lagrangian function of the optimization problem stated in Equation (21). If the optimization problem is intended to be a minimization problem, it should be explicitly mentioned to avoid any ambiguity or confusion.
	\reply {In the revised manuscript, we have simplified many notations and rewritten (40) properly. Problem (21) is a maximization problem and the sign of the Lagrangian function has been corrected.}

	\point The expression of the primary achievable rate in eq. (37) seems to be wrong since the primary message is also included in the cascaded channel and cannot be ignored.
	\reply {We appreciate your careful reading. The primary message is indeed included in the scattered component. However, since the user decodes both links independently, the backscatter uncertainty cannot be eliminated when decoding the primary link. That is to say, the scattered component should be treated as interference with average power $\mathbb{E}\bigl\{\sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_k \boldsymbol{w}s[n]\bigr\} = \sum_{k} \lvert \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} \boldsymbol{w} \rvert^2$. The same reasoning was considered in \cite{Long2020a}. We have revised the corresponding footnote to justify this.}

	\point In Figure 7, it is unclear why the KKT, PGD, and BCD algorithms are plotted separately since the BCD algorithm encompasses both the KKT and PGD algorithms. If there is a specific reason for displaying them separately, please provide a justification to enhance the understanding of the reader. Additionally, please provide the unit for the weighted sum rate axis.
	\reply {In company with BCD, we also plotted the convergence results of KKT and PGD algorithms to show how much performance gain is provided by solving each subproblem (over the initialization method specified in the previous paragraph). The unit of the y axis is bit/s/Hz and we have added this in the revised manuscript.}

	\point Is the decoding of the backscatter information based on an energy detector reliable in this case, especially with high-order modulation or large number of scatter nodes? In other words, can the user in practice decode the backscatter information especially in the presence of the noise and signal variations due to the fading or the interference, etc.?
	\reply {The energy detector is a simple and practical method for backscatter information decoding, and its reliability is improved by the adaptive input distribution and thresholding design. With high-order modulation or large number of scatter nodes, its performance can be further enhanced by increasing the spreading factor $N$ or using stronger error correction codes with lower code-rate. In practice, the user can decode backscatter nodes ranging from few to hundreds of meters in the presence of noise and interference, and the backscatter throughput can reach few Kbps to tens of Mbps. We have added a footnote on this, and the reviewer is referred to Table 2 in \cite{Wu2022e} for more details.}

	\point In Section III, the aim is to characterize the primary-total-backscatter rate region where the user retrieves both the backscatter information and the primary message. Why is the weighted sum mutual information maximized with respect to only the input distribution of the backscatter information and not jointly with respect to the input distribution of the primary message `s' as well? In other words, is the choice of $s \sim \mathcal{CN} (0,1)$ optimal?
	\reply {This is indeed a very good question. The choice of $s \sim \mathcal{CN} (0,1)$ is generally not optimal for weighted sum mutual information. To the best of our knowledge, the optimal primary distribution in this case remains an open issue, but some insights were given in those papers:
		\begin{itemize}
			\item If there is no direct link and the node is with \emph{average} power constraint, then the optimal distribution for \emph{weighted sum mutual information} requires a time-sharing between those modes \cite{Pillai2011}:
			\begin{itemize}
				\item $s$ follows zero-mean Gaussian distribution and $x$ is a constant;
				\item $x$ follows zero-mean Gaussian distribution and $s$ is a constant;
				\item $s$ and $x$ equal 0.
			\end{itemize}
			\item If there is no direct link and the node is with \emph{peak} power constraint, then the optimal distribution for \emph{sum mutual information} is not unique, for example \cite{Zhang2023} (came out after our submission):
			\begin{itemize}
				\item $s$ follows zero-mean Gaussian distribution and $x$ is a constant;
				\item $s$ follows Rayleigh distribution and $x$ follows uniform distribution on $[0, 2 \pi]$;
			\end{itemize}
		\end{itemize}
		Despite not optimal, the assumption of Gaussian primary source and finite-support backscatter source are relatively practical and widely adopted in relevant literatures. We have added Remark 3 on this, but intentionally avoided the discussion of optimal distribution, since it is beyond the scope of this paper and may confuse the readers. We appreciate your understanding in this matter.}

	\point In Section IV, Subsection B, it is stated that the RIS always ensures constructive superposition of direct and scattered components which is not true. The RIS can also add those components destructively which could be used for example to enhance the physical layer security of a system in the presence of an eavesdropper.
	\reply {We apologize for the oversight. It has been modified to ``Here, RIS ensures constructive superposition of direct and scattered components.''}
\end{reviewer}

### 2023-07-31 main.tex
:speech_balloon: Add new plots

		\begin{remark}
			We have assumed Gaussian codebook for the primary source and finite support for the backscatter nodes.
			The proposed framework is extendable to non-Gaussian primary source, and the conditional \gls{pdf} \eqref{eq:energy_distribution} can be approximated using \gls{clt} for large $N$.
		\end{remark}

		The user first jointly decodes the backscatter message of all nodes using a low-complexity energy detector.
		The energy detector formulates a \gls{dmtc} of size $M^K \times M^K$.

@@

		\begin{figure}[!t]
			\centering
			\resizebox{0.8\columnwidth}{!}{
				\input{assets/illustration/energy_distribution.tex}
			}
			\caption{
				\gls{pdf} of receive energy per backscatter block conditioned on different reflection state.
			}
			\label{fg:energy_distribution}
		\end{figure}

@@

\begin{section}{Simulation Results}
	In this section, we provide numerical results to evaluate the proposed input distribution, active beamforming, and backscatter decision designs for the considered RIScatter network.
	We assume the AP-user distance is \SI{10}{\meter} and at least one RIScatter nodes are randomly dropped in a disk centered at the user with radius $r$.
	% The \gls{ap} is with an average transmit power budget $P=\SI{36}{\dBm}$, all nodes employs $M$-\gls{qam} with amplitude scattering ratio $\alpha=0.5$, and the user is with average noise power $\sigma_v^2=\SI{-40}{\dBm}$.
	The \gls{ap} is with an average transmit power budget $P=\SI{36}{\dBm}$ and all nodes employs $M$-\gls{qam} with $\alpha=0.5$.
	For all channels involved, we consider a distance-dependent path loss model
	\begin{equation}
		L(d) = L_0 \biggl(\frac{d_0}{d}\biggr)^\gamma,
	\end{equation}
	together with a Rician fading model
	\begin{equation}
		\boldsymbol{H} = \sqrt{\frac{\kappa}{1+\kappa}} \bar{\boldsymbol{H}} + \sqrt{\frac{1}{1+\kappa}} \tilde{\boldsymbol{H}},
	\end{equation}
	where $d$ is the transmission distance, $L_0=-\SI{30}{\dB}$ is the reference path loss at $d_0=\SI{1}{\meter}$, $\kappa$ is the Rician K-factor, $\bar{\boldsymbol{H}}$ is the deterministic line-of-sight component with unit-magnitude entries, and $\tilde{\boldsymbol{H}}$ is the Rayleigh fading component with standard \gls{iid} \gls{cscg} entries.
	We choose $\gamma_{\text{D}}=2.6$, $\gamma_{\text{F}}=2.4$, $\gamma_{\text{B}}=2$, and $\kappa_{\text{D}}=\kappa_{\text{F}}=\kappa_{\text{B}}=5$ for direct, forward and backward links.
	The finite decision threshold domain $\mathcal{T}$ is obtained by $b$-bit uniform discretization over the critical interval defined by the confidence bounds of edge hypotheses (i.e., lower bound of $x_1$ and upper bound of $x_L$) with confidence $1-\varepsilon$.
	We set $b=9$ and $\varepsilon=\num{e-3}$.
	All achievable rate points/regions are averaged over \num{e3} channel realizations.

	\begin{subsection}{Evaluation of Proposed Algorithms}
		\begin{subsubsection}{Initialization}
			To characterize each achievable rate region, we progressively obtain all boundary points by successively increasing the primary \gls{qos} $\rho$ and solving problem \eqref{op:weighted_sum_rate}.
			For $\rho=0$ where the backscatter link is prioritized, we initialize Algorithm \ref{al:input_distribution} and \ref{al:active_beamforming} by uniform input distribution and \gls{mrt} towards the sum cascaded channel $\sum_{k} \boldsymbol{h}_{\text{C},k}^\mathsf{H}$, respectively.
			At the following points, both algorithms are initialized by the final solutions at the previous point.
		\end{subsubsection}

		\begin{subsubsection}{Convergence}
			\begin{figure}[!t]
				\centering
				\resizebox{0.75\columnwidth}{!}{
					\input{assets/simulation/wsr_convergence.tex}
				}
				\caption{Typical convergence curves at $\rho=0$ for $Q=4$, $K=8$, $M=2$, $N=20$, $\sigma_v^2=\SI{-40}{\dBm}$ and $r=\SI{2}{\meter}$.}
				\label{fg:wsr_convergence}
			\end{figure}
			The \gls{bcd} algorithm is convergent for the optimization problem \eqref{op:weighted_sum_rate}, since the input distribution and active beamforming subproblems converge and the thresholding subproblem attains global optimality.
			In Fig.~\ref{fg:wsr_convergence}, we plot the weighted sum of primary and total backscatter rates at $\rho=0$ for \gls{kkt}, \gls{pga} and \gls{bcd} algorithms on the first call.
			For $K=8$ and $M=2$, Algorithm \ref{al:input_distribution} typically takes around \num{100} fast iterations to converge to the \gls{kkt} input distribution.
			For $Q=4$, around \num{10} iterations are required for Algorithm \ref{al:active_beamforming} to converge.
			Overall, the initial \gls{bcd} algorithm requires at most \num{5} iterations to converge.
			At the following points (not presented here), the convergence of all three algorithms are much faster thanks to the progressive initialization.
			We conclude the proposed algorithms have good convergence performances.
		\end{subsubsection}
	\end{subsection}

	\begin{subsection}{Comparison of Scattering Applications}
		On top of the setup in Fig.~\ref{fg:riscatter_network}, we consider RIScatter and the following benchmark applications:
		\begin{itemize}
			\item \emph{Legacy:} Active transmission without scatterers.
			\item \emph{\gls{bbc}:} The primary symbol becomes deterministic $s[n]=1$ and the receive signal becomes
				y^{\text{\gls{bbc}}}[n] = \Bigl(\boldsymbol{h}_{\text{D}}^\mathsf{H} + \sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_k\Bigr) \boldsymbol{w} + v[n].
			The total backscatter rate of coherent decoding approaches $K \log M$ when $N$ is sufficiently large.
			% which follows non-zero mean complex Gaussian distribution $\mathcal{CN}\bigl((\boldsymbol{h}_{\text{D}}^\mathsf{H} + \sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_{m_k}) \boldsymbol{w},\sigma_v^2\bigr)$ under hypothesis $x_{m_{\mathcal{K}}}$.
			% The corresponding \gls{pdf} of receive energy over $N$ primary blocks is
			% \begin{equation}
			% 	f^{\text{\gls{bbc}}}(z|x_{m_{\mathcal{K}}}) = \frac{(z-\mu_{m_{\mathcal{K}}}^{\text{\gls{bbc}}})^{N-1} \exp \bigl(-(z-\mu_{m_{\mathcal{K}}}^{\text{\gls{bbc}}})/\sigma_v^2 \bigr)}{\sigma_v^{2N} (N-1)!},
			% 	\label{eq:energy_distribution_bbc}
			% \end{equation}
			% where $\mu_{m_{\mathcal{K}}}^{\text{\gls{bbc}}} \triangleq N \bigl\lvert \bigl(\boldsymbol{h}_{\text{D}}^\mathsf{H} + \sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_{m_k}\bigr) \boldsymbol{w} \bigr\rvert^2$.
			% The \gls{ml} decision threshold is derived as, $\forall l \in \mathcal{L} \setminus \{L\}$,
			% \begin{equation}
			% 	t_l^{\text{\gls{bbc}}} = \frac{\mu_{l-1}^{\text{\gls{bbc}}} \exp \bigl((\mu_{l-1}^{\text{\gls{bbc}}}-\mu_{l}^{\text{\gls{bbc}}})/\sigma_v^2 (N-1)\bigr) - \mu_{l}^{\text{\gls{bbc}}}}{\exp \bigl((\mu_{l-1}^{\text{\gls{bbc}}}-\mu_{l}^{\text{\gls{bbc}}})/\sigma_v^2 (N-1)\bigr) - 1}.
			% 	\label{eq:detection_threshold_ml_bbc}
			% \end{equation}
			\item \emph{\gls{ambc}:} The user decodes both links independently and semi-coherently by treating the other as interference.
			The primary achievable rate is approximately\footnote{To provide a preliminary benchmark, we regard the scattered component $\sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_k \boldsymbol{w}s[n]$ as \gls{awgn} during primary decoding \cite{Long2020a}.}
				I_{\text{P}}^{\text{\gls{ambc}}}(s;y) \approx \log \Bigl(1 + \frac{\lvert\boldsymbol{h}_{\text{D}}^\mathsf{H}\boldsymbol{w}\rvert^2}{\sum_{k}\lvert \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} \boldsymbol{w}\rvert^2+\sigma_v^2}\Bigr),
			while the total backscatter rate follows \eqref{eq:backscatter_mutual_information} with uniform input distribution.
			\item \emph{\gls{sr}:} For a sufficiently large $N$, the average primary rate under semi-coherent detection asymptotically approaches \eqref{eq:primary_mutual_information} with uniform input distribution \cite{Long2020a}.
			When $s[n]$ is successfully decoded and the direct interference $\boldsymbol{h}_{\text{D}}^\mathsf{H} \boldsymbol{w} s[n]$ is perfectly cancelled, the intermediate signal is
				\hat{y}^{\text{\gls{sr}}}[n] = \sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_k \boldsymbol{w} s[n] + v[n].
				\label{eq:intermediate_signal}
			The total backscatter rate of coherent decoding approaches $K \log M$ when $N$ is sufficiently large.
			% which only involves noise uncertainty under hypothesis $x_{m_{\mathcal{K}}}$.
			% During backscatter detection, the primary symbols $s[1],\ldots,s[N]$ function as a spreading code and the receiver performs \gls{mrc} on $\hat{y}^{\text{\gls{sr}}}[n]$ over $N$ primary blocks.
			% This essentially formulates a \gls{dcmc} and the total backscatter achievable rate is maximized with equiprobable inputs as \cite{Ng2006}
			% \begin{equation}
			% 	I_{\text{B}}(x_{\mathcal{K}};\hat{y}^{\text{\gls{sr}}}) = K \log M - \frac{\iota}{M^K},
			% \end{equation}
			% where $\iota \triangleq \sum_{m_{\mathcal{K}}} \mathbb{E}_{\hat{v}} \log \sum_{m_{\mathcal{K}}'} \exp ( - {\lvert x_{m_{\mathcal{K}}} - x_{m_{\mathcal{K}}'} + \hat{v} \rvert^2}/{2 \sigma^2} )$ and $\hat{v} \sim \mathcal{CN}(0,\sigma_v^2/N)$.
			% For a sufficiently large $N$, $\iota$ is negligible and the total backscatter rate approaches $K \log M$.
			\item \emph{\gls{ris}:} The reflection pattern is deterministic and the total backscatter rate is zero.
			The primary achievable rate is a special case of \eqref{eq:primary_mutual_information}
				I_{\text{P}}^{\text{\gls{ris}}}(s;y|x_{\mathcal{K}}) = I_{\text{P}}(s;y|x_{m_{\mathcal{K}}^{\star}}) = \log \Bigl(1 + \frac{\lvert \boldsymbol{h}^\mathsf{H}(x_{m_{\mathcal{K}}^{\star}}) \boldsymbol{w} \rvert^2}{\sigma_v^2}\Bigr),
			where $m_{\mathcal{K}}^{\star} = \arg \max_{m_{\mathcal{K}}} I_{\text{P}}^{\text{\gls{ris}}}(s;y|x_{m_{\mathcal{K}}})$.
		\end{itemize}
		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/simulation/region_comparison.tex}
			}
			\caption{Typical achievable rate region/points of scattering applications for $Q=1$, $K=1$, $M=4$, $N=\num{e3}$, $\sigma_v^2=\SI{-40}{\dBm}$ and $r=\SI{2}{\meter}$.}
			\label{fg:region_comparison}
		\end{figure}

		Fig.~\ref{fg:region_comparison} compares the typical achievable rate region/points of RIScatter and the strategies above.
		% \emph{First,} we observe both \gls{bbc} and \gls{sr} achieves when the spreading factor $N$ is sufficiently large.
		\emph{First,} we observe \gls{bbc} and \gls{sr} achieve the best backscatter performance thanks to coherent decoding.
		% However, \gls{sr} requires $N$ \gls{sic} followed by a \gls
		% However, \gls{sr} requires $N$ re-encoding, precoding, subtraction together with a time-domain \gls{mrc} to decode each backscatter symbol and the cost can be prohibitive.
		For \gls{sr}, this comes with the cost of $N$ re-encoding, precoding, subtraction together with a time-domain \gls{mrc} per backscatter symbol.
		% However, the primary rate benefit
		% For \gls{bbc} with coherent energy detection, the conditional \gls{pdf} of receive energy \eqref{eq:energy_distribution_bbc} is more skewed at a large $N$, such that the equivalent \gls{dmmac} \eqref{eq:dmmac} has lower error probabilities.
		% For \gls{sic}-based \gls{sr}, the effective backscatter \gls{snr} is increased by $N$ times and the penalty term $\iota$ becomes insignificant.
		% However, the primary rate is only maintained at a very large spreading factor, which significantly constrains the backscatter \emph{throughput}.
		\emph{Second,} the average primary rate slightly decreases/increases in the presence of a \gls{ambc}/\gls{ris} node and the benefit of \gls{sr} is not obvious.
		This is because the cascaded channel is usually orders of magnitude weaker than the direct channel.
		\gls{ris} always ensures constructive superposition of direct and scattered components while \gls{sr} only creates a quasi-static rich-scattering environment that marginally enhances the average primary rate.
		When $N$ is moderate, the randomly scattered signals should be modelled as interference rather than multipath components and the \gls{sr} point should move towards the \gls{ambc} point.
		\emph{Third,} RIScatter enables a flexible primary-backscatter tradeoff with adaptive input distribution design.
		In terms of maximum primary achievable rate, RIScatter coincides with \gls{ris} and outperforms the others using deterministic reflection pattern.
		On the other hand, for a large $N$, the maximum backscatter achievable rate of RIScatter is lower than \gls{bbc} and \gls{sr} but higher than \gls{ambc}.
		This is because both RIScatter and \gls{ambc} employ semi-coherent energy detection, and the adaptive channel coding of RIScatter provides better performance than equiprobable inputs of \gls{ambc} especially at the low \gls{snr} of double fading.
		When multiple antenna is available at the \gls{ap}, active beamforming can be optimized for RIScatter nodes and the backscatter rate gain over \gls{ambc} can be even larger.
	\end{subsection}

>>>>

		\begin{remark}
			We have assumed Gaussian codebook for the primary source and finite support for the backscatter nodes, since they are relatively practical and widely adopted in relevant literatures.
			The proposed framework is extendable to non-Gaussian primary source, and the conditional \gls{pdf} \eqref{eq:energy_distribution} can be approximated using \gls{clt} for large $N$.
		\end{remark}

		The user first jointly decodes the backscatter message of all nodes using a low-complexity energy detector.\footnote{The reliability of the energy detector is improved by the adaptive input distribution and thresholding design. With high-order modulation or large number of scatter nodes, the reliability can be further enhanced by increasing the spreading factor or using error correction codes with lower code-rate. In practice, users can decode backscatter nodes ranging from a few to hundreds of meters in the presence of noise and interference, and the backscatter throughput can reach few Kbps to tens of Mbps \cite{Wu2022e}.}
		The energy detector formulates a \gls{dmtc} of size $M^K \times M^K$.

@@

		\begin{figure}[!t]
			\centering
			\resizebox{0.8\columnwidth}{!}{
				\input{assets/illustration/energy_distribution.tex}
			}
			\caption{
				\gls{pdf} of the receive energy per backscatter block conditioned on different reflection state.
			}
			\label{fg:energy_distribution}
		\end{figure}

@@

\begin{section}{Simulation Results}
	In this section, we provide numerical results to evaluate the proposed input distribution, active beamforming, and backscatter decision algorithms.
	We assume the AP-user distance is \SI{10}{\meter} and at least one RIScatter nodes are randomly dropped in a disk centered at the user with radius $r$.
	The \gls{ap} is with an average transmit power budget $P=\SI{36}{\dBm}$ and all nodes employs $M$-\gls{qam} with $\alpha=0.5$.
	For all channels involved, we consider a distance-dependent path loss model
	\begin{equation}
		L(d) = L_0 \biggl(\frac{d_0}{d}\biggr)^\gamma,
	\end{equation}
	together with a Rician fading model
	\begin{equation}
		\boldsymbol{H} = \sqrt{\frac{\kappa}{1+\kappa}} \bar{\boldsymbol{H}} + \sqrt{\frac{1}{1+\kappa}} \tilde{\boldsymbol{H}},
	\end{equation}
	where $d$ is the transmission distance, $L_0=-\SI{30}{\dB}$ is the reference path loss at $d_0=\SI{1}{\meter}$, $\kappa$ is the Rician K-factor, $\bar{\boldsymbol{H}}$ is the deterministic line-of-sight component with unit-magnitude entries, and $\tilde{\boldsymbol{H}}$ is the Rayleigh fading component with standard \gls{iid} \gls{cscg} entries.
	We choose $\gamma_{\text{D}}=2.6$, $\gamma_{\text{F}}=2.4$, $\gamma_{\text{B}}=2$, and $\kappa_{\text{D}}=\kappa_{\text{F}}=\kappa_{\text{B}}=5$ for direct, forward and backward links.
	The finite decision threshold domain $\mathcal{T}$ is obtained by $b$-bit uniform discretization over the critical interval defined by the confidence bounds of edge hypotheses (i.e., lower bound of $x_1$ and upper bound of $x_L$) with confidence $1-\varepsilon$.
	We set $b=9$ and $\varepsilon=\num{e-3}$.
	All achievable rate regions are averaged over \num{e3} channel realizations.

	\begin{subsection}{Evaluation of Proposed Algorithms}
		\begin{subsubsection}{Initialization}
			To characterize the achievable rate region, we progressively obtain all boundary points by successively increasing $\rho$ and solving problem \eqref{op:weighted_sum_rate}.
			For $\rho=0$ where the backscatter link is prioritized, we initialize Algorithm \ref{al:input_distribution} and \ref{al:active_beamforming} by uniform input distribution and \gls{mrt} towards the sum cascaded channel $\sum_{k} \boldsymbol{h}_{\text{C},k}^\mathsf{H}$, respectively.
			At the following points, both algorithms are initialized by the solutions at the previous point.
		\end{subsubsection}

		\begin{subsubsection}{Convergence}
			\begin{figure}[!t]
				\centering
				\resizebox{0.75\columnwidth}{!}{
					\input{assets/simulation/wsr_convergence.tex}
				}
				\caption{Typical convergence curves at $\rho=0$ for $Q=4$, $K=8$, $M=2$, $N=20$, $\sigma_v^2=\SI{-40}{\dBm}$ and $r=\SI{2}{\meter}$.}
				\label{fg:wsr_convergence}
			\end{figure}
			The \gls{bcd} algorithm is convergent for problem \eqref{op:weighted_sum_rate} since the input distribution and active beamforming subproblems converge and the thresholding subproblem attains global optimality.
			In company with \gls{bcd}, we also plotted the convergence results of \gls{kkt} and \gls{pga} algorithms in Fig.~\ref{fg:wsr_convergence} to show how much performance is gained by solving each subproblem.
			It is observed that Algorithm \ref{al:input_distribution} and \ref{al:active_beamforming} take around \num{100} and \num{10} iterations to converge, respectively.
			Overall, the \gls{bcd} algorithm requires at most \num{5} iterations to converge.
			As $\rho$ increases (not presented here), the convergence of all three algorithms are much faster thanks to the progressive initialization.
		\end{subsubsection}
	\end{subsection}

	\begin{subsection}{Comparison of Scattering Applications}
		On top of the setup in Fig.~\ref{fg:riscatter_network}, we consider RIScatter and the following benchmark applications:
		\begin{itemize}
			\item \emph{Legacy:} Active transmission without scatterers.
			\item \emph{\gls{bbc}:} The primary symbol becomes deterministic $s[n]=1$ and the receive signal becomes
			\begin{equation}
				y^{\text{\gls{bbc}}}[n] = \Bigl(\boldsymbol{h}_{\text{D}}^\mathsf{H} + \sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_k\Bigr) \boldsymbol{w} + v[n].
			\end{equation}
			The total backscatter rate of coherent decoding approaches $K \log M$ when $N$ is sufficiently large.
			\item \emph{\gls{ambc}:} The user decodes both links independently and semi-coherently by treating the other as interference.
			The primary achievable rate is approximately\footnote{The scattered component is treated as interference with average power $\mathbb{E}\bigl\{\sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_k \boldsymbol{w}s[n]\bigr\} = \sum_{k} \lvert \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} \boldsymbol{w} \rvert^2$ \cite{Long2020a}.}
			\begin{equation}
				I_{\text{P}}^{\text{\gls{ambc}}}(s;y) \approx \log \Bigl(1 + \frac{\lvert\boldsymbol{h}_{\text{D}}^\mathsf{H}\boldsymbol{w}\rvert^2}{\sum_{k}\lvert \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} \boldsymbol{w}\rvert^2+\sigma_v^2}\Bigr),
			\end{equation}
			while the total backscatter rate follows \eqref{eq:backscatter_mutual_information} with uniform input distribution.
			\item \emph{\gls{sr}:} For a sufficiently large $N$, the average primary rate under semi-coherent detection asymptotically approaches \eqref{eq:primary_mutual_information} with uniform input distribution \cite{Long2020a}.
			When $s[n]$ is successfully decoded and the direct interference $\boldsymbol{h}_{\text{D}}^\mathsf{H} \boldsymbol{w} s[n]$ is perfectly cancelled, the intermediate signal is
			\begin{equation}
				\hat{y}^{\text{\gls{sr}}}[n] = \sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_k \boldsymbol{w} s[n] + v[n].
				\label{eq:intermediate_signal}
			\end{equation}
			The total backscatter rate of coherent decoding approaches $K \log M$ when $N$ is sufficiently large.
			\item \emph{\gls{ris}:} The reflection pattern is deterministic and the total backscatter rate is zero.
			The primary achievable rate is a special case of \eqref{eq:primary_mutual_information}
			\begin{equation}
				I_{\text{P}}^{\text{\gls{ris}}}(s;y|x_{\mathcal{K}}) = I_{\text{P}}(s;y|x_{m_{\mathcal{K}}^{\star}}) = \log \Bigl(1 + \frac{\lvert \boldsymbol{h}^\mathsf{H}(x_{m_{\mathcal{K}}^{\star}}) \boldsymbol{w} \rvert^2}{\sigma_v^2}\Bigr),
			\end{equation}
			where $m_{\mathcal{K}}^{\star} = \arg \max_{m_{\mathcal{K}}} I_{\text{P}}^{\text{\gls{ris}}}(s;y|x_{m_{\mathcal{K}}})$.
		\end{itemize}
		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/simulation/region_comparison.tex}
			}
			\caption{Typical achievable rate region/points of scattering applications for $Q=1$, $K=1$, $M=4$, $N=\num{e3}$, $\sigma_v^2=\SI{-40}{\dBm}$ and $r=\SI{2}{\meter}$.}
			\label{fg:region_comparison}
		\end{figure}

		Fig.~\ref{fg:region_comparison} compares the typical achievable rate region/points of RIScatter and those strategies.
		\emph{First,} we observe \gls{bbc} and \gls{sr} achieve the best backscatter performance thanks to coherent decoding.
		For \gls{sr}, this comes with the cost of $N$ re-encoding, precoding, subtraction together with a time-domain \gls{mrc} per backscatter symbol.
		\emph{Second,} the average primary rate slightly decreases/increases in the presence of a \gls{ambc}/\gls{ris} node and the benefit of \gls{sr} is not obvious.
		This is because the cascaded channel is usually orders of magnitude weaker than the direct channel.
		Here, \gls{ris} ensures constructive superposition of direct and scattered components while \gls{sr} only creates a quasi-static rich-scattering environment that marginally enhances the average primary rate.
		When $N$ is moderate, the randomly scattered signals should be modelled as interference rather than multipath components and the \gls{sr} point should move towards the \gls{ambc} point.
		\emph{Third,} RIScatter enables a flexible primary-backscatter tradeoff with adaptive input distribution design.
		In terms of maximum primary achievable rate, RIScatter coincides with \gls{ris} and outperforms the others using deterministic reflection pattern.
		On the other hand, for a large $N$, the maximum backscatter achievable rate of RIScatter is lower than \gls{bbc} and \gls{sr} but higher than \gls{ambc}.
		This is because both RIScatter and \gls{ambc} employ semi-coherent energy detection, and the adaptive channel coding of RIScatter outperforms the equiprobable inputs of \gls{ambc}, especially at the low backscatter \gls{snr} in the double fading scenario.
		When multiple antenna is available at the \gls{ap}, active beamforming can be optimized for RIScatter nodes and the backscatter rate gain over \gls{ambc} is more prominent.
	\end{subsection}

	\begin{subsection}{Input Distribution under Different \gls{qos}}
		\begin{figure}[!t]
			\centering
			\resizebox{0.65\columnwidth}{!}{
				\input{assets/simulation/distribution_weights.tex}
			}
			\caption{Typical RIScatter reflection state distribution at different $\rho$ for $Q=1$, $K=1$, $M=4$, $N=20$, $\sigma_v^2=\SI{-40}{\dBm}$ and $r=\SI{2}{\meter}$.}
			\label{fg:distribution_weights}
		\end{figure}
		The objective is to demonstrate RIScatter nodes can leverage \gls{csi}- and \gls{qos}-adaptive input distribution design to balance backscatter modulation and passive beamforming.
		For one RIScatter node with $M=4$, we evaluate the \gls{kkt} input distribution at different \gls{qos} and present the result in Fig.~\ref{fg:distribution_weights}.
		At $\rho=0$ where the backscatter performance is prioritized, the optimal input distribution is \num{0} on two states and nearly uniform on the other two.
		This is inline with Shannon's observation that binary antipodal inputs is good enough for channel capacity at the low \gls{snr} \cite{Shannon1948}.
		Due to the weak scattered signal, the conditional energy \gls{pdf} under different hypotheses can be closely spaced as in Fig.~\ref{fg:energy_distribution}.
		The extreme states producing the lowest/highest energy are always assigned with non-zero probability, while the middles cannot provide enough energy difference and thus end up unused.
		At $\rho=1$ where the primary performance is prioritized, the optimal input distribution is \num{1} at the state that maximizes the primary \gls{snr} and \num{0} at the others.
		That is, the reflection pattern becomes deterministic and the RIScatter node boils down to a discrete \gls{ris} element.
		Increasing $\rho$ from \num{0} to \num{1} provides a smooth transition from backscatter modulation to passive beamforming, suggesting RIScatter unifies \gls{bc} and \gls{ris} from a probabilistic perspective.
	\end{subsection}

	\begin{subsection}{Rate Region by Different Schemes}
		\begin{figure}[!t]
			\centering
			\subfloat[Input Distribution, $Q=1$\label{fg:region_distribution}]{
				\resizebox{0.6\columnwidth}{!}{
					\input{assets/simulation/region_distribution.tex}
				}
			}
			\\
			\subfloat[Active Beamforming, $Q=4$\label{fg:region_beamforming}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/region_beamforming.tex}
				}
			}
			\subfloat[Decision Threshold, $Q=4$\label{fg:region_threshold}]{
				\resizebox{0.48\columnwidth}{!}{
					\input{assets/simulation/region_threshold.tex}
				}
			}
			\caption{
				Average primary-total-backscatter rate regions by different input distribution, active beamforming, and decision threshold schemes for $K=2$, $M=4$, $N=20$, $\sigma_v^2=\SI{-40}{\dBm}$ and $r=\SI{2}{\meter}$.
			}
		\end{figure}
		\begin{subsubsection}{Input Distribution}
			We compare these input distribution designs for problem \eqref{op:input_distribution}:
			\begin{itemize}
				\item \emph{Cooperation:} Joint encoding using a joint probability array $p(x_{m_{\mathcal{K}}})$ with $M^K$ entries by Algorithm \ref{al:input_distribution};
				\item \emph{Exhaustion:} Exhaustive search over the $M$-dimensional probability simplex with resolution $\Delta p = \num{e-2}$;
				\item \emph{\gls{kkt}:} The \gls{kkt} solution by Algorithm \ref{al:input_distribution};
				\item \emph{Equiprobable:} Uniform input distribution.
			\end{itemize}
			We also consider these independent distribution recovery methods from the joint probability array:
			\begin{itemize}
				\item \emph{Marginalization:} Marginal probability distributions;
				\item \emph{Decomposition:} Normalized rank-\num{1} \gls{cp} decomposed tensors by \texttt{Tensor Toolbox} \cite{Bader2022};
				\item \emph{Randomization:} Gaussian randomization with the guidance of correlation matrix \cite{Calvo2010}.
			\end{itemize}
